---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# stability

<!-- badges: start -->
<!-- badges: end -->

The goal of stability is to analyse stability data to determine product expiry date based on pre-defined upper/lower limits in the drug development process. Observed stability results and expiry date based on the chosen stability model are visualised by time (months) and batch. 

## Installation

You can install the development version of stability from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("joanna-ling/stability")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
#### Load Required Packages ####
if (!require("pacman")) install.packages("pacman")

suppressMessages(
  pacman::p_load(
    tidyverse,
    ggplot2,
    readxl,
    janitor,
    stringr,
    ggpubr,
    stability
  )
)

# Test using CSL312 sFP 36 months example data
data <- read_excel("C:/Users/U0152563/OneDrive - CSL Behring/Desktop/R learning/stability/Stability_Examples.xlsx", sheet = "CLS312_sFP_36m") %>% 
  clean_names() %>% 
  mutate(time = time_months,
         batch = lot) %>% 
  subset(!(temperature == "2 to 8°C" & time == 36)) %>% 
  # This part is manually removing values that are different from the report... 
  mutate(flag = case_when(quality_attribute == "SE HPLC Monomer" & result_value == 97.2 & time == 3 & batch == "P100213605" ~ 1,
                          quality_attribute == "SE HPLC Monomer" & result_value == 97.5 & time == 3 & batch == "P100229828" ~ 1,
                          quality_attribute == "SE HPLC Monomer" & result_value == 97.5 & time == 9 & batch == "P100229828" ~ 1,
                          quality_attribute == "SE HPLC HMWS" & result_value == 3.0 & time == 3 & batch == "P100213605" ~ 1,
                          quality_attribute == "SE HPLC HMWS" & result_value == 2.1 & time == 3 & batch == "P100229828" ~ 1,
                          quality_attribute == "SE HPLC HMWS" & result_value == 2.5 & time == 9 & batch == "P100229828" ~ 1,
                          TRUE ~ 0 )) %>% 
  subset(flag == 0)

data_dist <- data[,c("quality_attribute","units","lower_limit","upper_limit")] %>% 
  distinct() %>% 
  subset(!is.na(lower_limit) | !is.na(upper_limit)) %>% 
  mutate(label = case_when(units == "N/A" ~ paste0(quality_attribute),
                           TRUE ~ paste0(quality_attribute," (",units,")")))

data_transf <- data %>% 
  mutate(time = case_when(quality_attribute == "SE HPLC Monomer" ~ sqrt(time),
                          quality_attribute == "SE HPLC HMWS" ~ sqrt(time),
                          quality_attribute == "CE reducing" ~ sqrt(time),
                          quality_attribute == "CE non reducing LMWS" ~ sqrt(time),
                          TRUE ~ time),
         backtransform = case_when(quality_attribute == "SE HPLC Monomer" ~ "Time^2",
                                   quality_attribute == "SE HPLC HMWS" ~ "Time^2",
                                   quality_attribute == "CE reducing" ~ "Time^2",
                                   quality_attribute == "CE non reducing LMWS" ~ "Time^2",
                                   TRUE ~ "Time"))


# Set colours using all batches in dataset
colours <- c("#FC1921","#0E56A5","#F5C017","#975DA2","#00A28A",
             "#DA2877","#F06125","#03B3BE","#C6D92D","#572A7B")
colours_named <- setNames(object = colours, nm = c(unique(data_transf$batch),"Common Intercept"))

qa <- data_dist$quality_attribute
lbl <- data_dist$label
ll <- data_dist$lower_limit
ul <- data_dist$upper_limit

# Initialise empty results list 
stab <- vector(mode = "list", length = length(qa))

for (i in 1:1) {
  
  data_qa <- subset(data_transf, quality_attribute == qa[i])
  
  bt <- unique(data_qa$backtransform)
  
  stab[[i]] <- stability(data = data_qa, outcome = "result_value", label = lbl[i],
                         lowerlimit = ll[i], upperlimit = ul[i], poolMSE = 0, 
                         backtransform = bt, colours_named = colours_named, debug = 0)
  
}

stab[[1]]

```
