---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# stability

<!-- badges: start -->
<!-- badges: end -->

The goal of stability is to analyse stability data to determine product expiry date based on pre-defined upper/lower limits in the drug development process. Observed stability results and expiry date based on the chosen stability model are visualised by time (months) and batch. 

## Installation

You can install the development version of stability from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("joanna-ling/stability")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}

# pak::pak("joanna-ling/stability")

#### Load Required Packages ####
if (!require("pacman")) install.packages("pacman")

suppressMessages(
  pacman::p_load(
    tidyverse,
    ggplot2,
    readxl,
    janitor,
    stringr,
    ggpubr,
    stability
  )
)

# Generate example data for a stability study
set.seed(873951)
example_data <- tibble(quality_attribute = c("SE HPLC HMWS", "pH", "CE Reducing Main Peak")) %>% 
  cross_join(tibble(time_months = c(0, 3, 6, 9, 12, 18, 24, 36))) %>%
  cross_join(tibble(batch = c("A", "B", "C"))) %>% 
  mutate(result_value = case_when(quality_attribute == "pH" ~ round(runif(n(), min = 5.9, max = 6.1),2),
                           quality_attribute == "SE HPLC HMWS" & time_months == 0 ~ round(rnorm(n(), mean = 1.2, sd = 0.05),2),
                           quality_attribute == "SE HPLC HMWS" & time_months == 3 ~ round(rnorm(n(), mean = 1.3, sd = 0.05),2),
                           quality_attribute == "SE HPLC HMWS" & time_months == 6 ~ round(rnorm(n(), mean = 1.5, sd = 0.05),2),
                           quality_attribute == "SE HPLC HMWS" & time_months == 9 ~ round(rnorm(n(), mean = 1.7, sd = 0.05),2),
                           quality_attribute == "SE HPLC HMWS" & time_months == 12 ~ round(rnorm(n(), mean = 1.9, sd = 0.05),2),
                           quality_attribute == "SE HPLC HMWS" & time_months == 18 ~ round(rnorm(n(), mean = 2, sd = 0.05),2),
                           quality_attribute == "SE HPLC HMWS" & time_months == 24 ~ round(rnorm(n(), mean = 2.1, sd = 0.05),2),
                           quality_attribute == "SE HPLC HMWS" & time_months == 36 ~ round(rnorm(n(), mean = 2.2, sd = 0.05),2),
                           quality_attribute == "CE Reducing Main Peak" & time_months == 0 ~ round(rnorm(n(), mean = 97, sd = 0.07),2),
                           quality_attribute == "CE Reducing Main Peak" & time_months == 3 ~ round(rnorm(n(), mean = 96.8, sd = 0.07),2),
                           quality_attribute == "CE Reducing Main Peak" & time_months == 6 ~ round(rnorm(n(), mean = 96.6, sd = 0.07),2),
                           quality_attribute == "CE Reducing Main Peak" & time_months == 9 ~ round(rnorm(n(), mean = 96.4, sd = 0.07),2),
                           quality_attribute == "CE Reducing Main Peak" & time_months == 12 ~ round(rnorm(n(), mean = 96.3, sd = 0.07),2),
                           quality_attribute == "CE Reducing Main Peak" & time_months == 18 ~ round(rnorm(n(), mean = 96.2, sd = 0.07),2),
                           quality_attribute == "CE Reducing Main Peak" & time_months == 24 ~ round(rnorm(n(), mean = 96.1, sd = 0.07),2),
                           quality_attribute == "CE Reducing Main Peak" & time_months == 36 ~ round(rnorm(n(), mean = 96.0, sd = 0.07),2),
                           TRUE ~ NA),
         lower_limit = case_when(quality_attribute == "pH" ~ 5.8,
                                 quality_attribute == "CE Reducing Main Peak" ~ 94.5,
                                 TRUE ~ NA),
         upper_limit = case_when(quality_attribute == "pH" ~ 6.2,
                                 quality_attribute == "SE HPLC HMWS" ~ 4.3,
                                 TRUE ~ NA),
         units = case_when(quality_attribute == "pH" ~ "",
                           quality_attribute == "SE HPLC HMWS" ~ "%",
                           quality_attribute == "CE Reducing Main Peak" ~ "%",
                           TRUE ~ NA),
         label = case_when(quality_attribute == "pH" ~ "pH",
                          TRUE ~ paste0(quality_attribute, " (", units, ")")),
         time = time_months)



# Get a dataset of distinct limits and generate clean labels 
data_dist <- example_data[,c("quality_attribute","units","label","lower_limit","upper_limit")] %>% 
  distinct() %>% 
  subset(!is.na(lower_limit) | !is.na(upper_limit)) 

# Transform time variable where needed and generate back transform variable - trial and error
data_transf <- example_data %>% 
  mutate(time = case_when(quality_attribute == "SE HPLC HMWS" ~ time^(1/2),
                          quality_attribute == "CE Reducing Main Peak" ~ time^(1/3),
                          TRUE ~ time),
         backtransform = case_when(quality_attribute == "SE HPLC HMWS" ~ "Time^2",
                                   quality_attribute == "CE Reducing Main Peak" ~ "Time^3",
                                   TRUE ~ "Time"))

# Set colours using all batches in dataset
colours <- c("#FC1921","#0E56A5","#F5C017","#975DA2","#00A28A",
             "#DA2877","#F06125","#03B3BE","#C6D92D","#572A7B")
colours_named <- setNames(object = colours, nm = c(unique(data_transf$batch),"Common Intercept"))

# Get vectors of quality attributes, labels and limits
qa <- data_dist$quality_attribute
lbl <- data_dist$label
ll <- data_dist$lower_limit
ul <- data_dist$upper_limit

# Initialise empty results list 
stab <- vector(mode = "list", length = length(qa))

# Loop over all quality attributes and save results to list
for (i in 1:length(qa)) {
  
  # Subset selected quality attribute
  data_qa <- subset(data_transf, quality_attribute == qa[i])
  
  # Get backtransformation string for the selected attribute
  bt <- unique(data_qa$backtransform)
  
  # Run stability function and save results to list
  stab[[i]] <- stability(data = data_qa, outcome = "result_value", label = lbl[i],
                         lowerlimit = ll[i], upperlimit = ul[i], poolMSE = 0, 
                         backtransform = bt, colours_named = colours_named, debug = 0)
  
}

# The results returned in the list contain: 
#   1. The chosen model as a number
#   2. The chosen model as a string 
#   3. A tibble of the estimated expiry dates and model fit with 90/95% confidence intervals for each model and batch
#   4. A tibble of the parameter estimates, standard errors, t value and p-value for each model 
#   5. The estimated expiry date from the chosen model as a number
#   6. The plot of observed stability data and predicted model fit with the appropriate confidence intervals up till the expiry date
#   7. A list of residual plots for models 1 to 4. 

stab[[1]]

```
